### 2024-05-28 presentation.tex
Initial commit.

\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{tikz}

\setmainlanguage{english}
\usetheme{cambridge}

\graphicspath{{imgs/}{svgs/}}

% TikZ libraries and styles
\usetikzlibrary{positioning,shapes.geometric,graphs,graphdrawing,quotes,fit}
\usegdlibrary{layered,trees}
\colorlet{GenericItem}{yellow!20!white}
\colorlet{GenericHighlight}{red!20!white}
\tikzset{Term/.style={draw,fill=yellow!20!white,font=\tiny\bfseries}}
\tikzset{Relation/.style={draw=blue!40!white,thick,arrows=<-}}
\tikzset{RelationLabel/.style={shape=circle,fill=white,draw=black,thin,font=\tiny\bfseries,inner sep=1pt}}
\tikzset{File/.style={draw,fill=yellow!20!white,font=\tiny\ttfamily}}
\tikzset{Entity/.style={draw,fill=yellow!20!white,font=\scriptsize}}
\tikzset{Process/.style={draw,shape=ellipse,fill=blue!20!white,font=\scriptsize}}

\title{(Re-)bridging the anatomy ontologies with SSSOM}

\author{Damien~Goutte-Gattat\inst{1}}
\institute{\inst{1}FlyBase group, Department of Physiology, Development and Neuroscience}
\date{International Conference on Biomedical Ontologies 2024, July 2024}

\hypersetup{pdfpagemode=UseNone,
  pdftitle={},
  pdfauthor={Damien Goutte-Gattat}}
\AtBeginSection{
  \begin{frame}{Outline}{}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{Introduction}

\begin{frame}
  \frametitle{An approach to a multi-species anatomy ontology}

  \begin{columns}
    \begin{column}{.5\textwidth}
      \begin{tikzpicture}[nodes={shape=circle,draw,fill=GenericItem}]
        \node (core) {Uberon};
        \node (fbbt) at (150:2cm) {FBbt};
        \node (zfa)  at (90:2cm)  {ZFA};
        \node (xao)  at (30:2cm)  {XAO};
        \node (ma)   at (210:2cm) {MA};
        \node (wbbt) at (270:2cm) {WBbt};
        \node (tx1) at  (310:2cm) {...};
        \node (tx2) at  (330:2cm) {...};
        \node (tx3) at  (350:2cm) {...};
        \draw (core) -- (fbbt) (core) -- (zfa)  (core) -- (xao)
              (core) -- (ma)   (core) -- (wbbt) (core) -- (tx1)
              (core) -- (tx2)  (core) -- (tx3);
      \end{tikzpicture}
    \end{column}
    \begin{column}{.45\textwidth}
      Strategy:
      \begin{itemize}
        \item Provide a taxon-neutral core framework
        \item Integrate existing taxon-specific ontologies
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Example: integration with FBbt}

  \only<1>{
    \begin{columns}[t]
      \begin{column}{.45\textwidth}
        Uberon hierarchy
        \begin{tikzpicture}
          \graph[layered layout,orient=down,nodes={Term},edge quotes mid,edges={Relation,nodes={RelationLabel}}] {
            anatomical entity <-
            anatomical structure <-
            circulatory organ <- {
              primary circulatory organ <- { heart, dorsal vessel heart },
              accessory circulatory organ <- lymph heart
            };
          };
        \end{tikzpicture}
      \end{column}
      \begin{column}{.45\textwidth}
        FBbt hierarchy
        \begin{tikzpicture}
          \graph[layered layout,orient=down,nodes={Term},edge quotes mid,edges={Relation,nodes={RelationLabel}}] {
            anatomical entity <-
            anatomical structure <-
            heart <-["P"] {
              aortic funnel, heart chamber, dorsal diaphragm
            };
          };
        \end{tikzpicture}
      \end{column}
    \end{columns}
  }
  \only<2>{
    \begin{columns}
      \begin{column}{.3\textwidth}
        Integrated hierarchy
      \end{column}
      \begin{column}{.6\textwidth}
        \begin{tikzpicture}
          \graph[layered layout,orient=down,nodes={Term},edge quotes mid,edges={Relation,nodes={RelationLabel}}] {
            anatomical entity <-
            anatomical structure <-
            circulatory organ <- {
              primary circulatory organ <- {
                heart,
                dorsal vessel heart <- "heart (drosophila)" <-["P"] {
                  aortic funnel, heart chamber, dorsal diaphragm
                }
              },
              accessory circulatory organ <- lymph heart
            };
          };
        \end{tikzpicture}
      \end{column}
    \end{columns}
  }
\end{frame}

\begin{frame}
  \frametitle{Integration requires cross-ontology mappings}

  \begin{block}{}
    Integrating the taxon-specific ontologies with the Uberon core framework
    requires to know where taxon-specific terms should be ``plugged'' in the
    core framework.

    E.g., we need to know that FBbt's \emph{heart} must be attached to Uberon's
    \emph{dorsal vessel heart}.
  \end{block}
\end{frame}

\section{Current system}

\begin{frame}
  \frametitle{Current solution: cross-references}

  \begin{block}{}\ttfamily
[Term]

id: UBERON:0015230

name: dorsal vessel heart

def: "The caudal, pulsatile region of the dorsal vessel of the arthropod circulatory system."

is\_a: UBERON:0007100 ! primary circulatory organ

xref: FBbt:00003154
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{From cross-references to bridge files}

  \begin{columns}
    \begin{column}{.3\textwidth}
      \begin{tikzpicture}
        \node [File] (uberon) at (0,0)     {uberon.obo};
        \node [File] (script) at (0,-1)    {make-bridge-ontology-from-xrefs.pl};
        \node [File] (bridge) at (0,-2)    {uberon-bridge-to-fbbt.obo};
        \node [File]          at (0, -2.5) {uberon-bridge-to-wbbt.obo};
        \node [File]          at (0, -3)   {uberon-bridge-to-ma.obo};
        \node [File]          at (0, -3.5) {uberon-bridge-to-....obo}; 
        \draw[->] (uberon) -- (script) (script) -- (bridge);
      \end{tikzpicture}
    \end{column}
    \begin{column}{.68\textwidth}
      \begin{block}{Bridge file}\ttfamily
[Term]

id: FBbt:00003154

property\_value: IAO:0000589 "heart (drosophila)" xsd:string

intersection\_of: UBERON:0015230

intersection\_of: part\_of NCBITaxon:7227
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Limitations of the current system}

  Limitations of cross-references to represent mappings
  \begin{itemize}
    \item Overloading of the \emph{oboInOwl:hasDbXref} annotation
    \item No metadata
    \item Mappings are embedded within the ontology and must be extracted
  \end{itemize}

  \bigskip
  Limitations of the bridging pipeline
  \begin{itemize}
    \item Highly Uberon-specific -- cannot be reused outside of Uberon
    \item Dependent on the OBO format
  \end{itemize}
\end{frame}

\section{New approach}

\begin{frame}
  \frametitle{Principles}

  \begin{enumerate}
    \item Clear separation between cross-ontology \emph{mappings} (COM) and cross-ontology \emph{bridges} (COB)
      \begin{itemize}
        \item A COB is merely the OWL materialisation of a COM
      \end{itemize}
    \item Represent COM using a standard format intended for that purpose
      \begin{itemize}
        \item allowing for mapping-specific metadata
        \item allowing for meaningful mapping predicates
      \end{itemize}
    \item Make it easy to derive COB from COM
      \begin{itemize}
        \item without requiring Uberon-specific tooling
      \end{itemize}
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{The SSSOM standard}

  ``Simple Standard for Sharing Ontological Mappings''
  \begin{itemize}
    \item A data model to represent semantic mappings and associated metadata
    \item A simple TSV-based format to store and exchange them
  \end{itemize}

  \begin{block}{SSSOM/TSV format}\tiny\ttfamily
    \begin{tabular}{lllll}
      \hline
      subject\_id   & subject\_label                & predicate\_id                 & object\_id      & mapping\_justification\\
      \hline
      FBbt:00000001 & organism                      & semapv:crossSpeciesExactMatch & UBERON:0000468  & semapv:ManualMappingCuration\\
      FBbt:00000002 & tagma                         & semapv:crossSpeciesExactMatch & UBERON:6000002  & semapv:ManualMappingCuration\\
      ...           & ...                           & ...                           & ...             & ...\\
      FBbt:00003154 & heart                         & semapv:crossSpeciesExactMatch & UBERON:0015230  & semapv:ManualMappingCuration\\
      ...           & ...                           &...                           & ...             & ...\\
      \hline
    \end{tabular}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{New mapping predicates for cross-species mappings}

  Additions to the \emph{Semantic Mapping Vocabulary} (SEMAPV)
  \begin{center}
    \begin{tikzpicture}
      \graph[tree layout,nodes={Term},edges={Relation}] {
        "skos:mappingRelation" <- {
          "semapv:isomorphicMatch" <- {
            "skos:exactMatch",
            "semapv:crossSpeciesExactMatch" [fill=GenericHighlight]
          },
          "semapv:nonIsomorphicMatch" <- {
            "skos:broadMatch" <- "semapv:crossSpeciesBroadMatch" [fill=GenericHighlight],
            "skos:narrowMatch" <- "semapv:crossSpeciesNarrowMatch" [fill=GenericHighlight]
          }
        };
      };
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}[t]
  \frametitle{From SSSOM sets to bridge files: SSSOM/T-OWL}

  A domain-specific language to transform SSSOM mappings into OWL axioms

  \only<1>{
    \begin{block}{Model of a mapping processing rule}
      \begin{center}
        \begin{tikzpicture}
          \node [text width=5cm,fill=GenericItem,draw] (filter) at (0,0)
                {\textbf{Filter}\\
                 Mapping $\rightarrow$ boolean\\
                 Decide whether the rule applies to a mapping};
          \node [text width=5cm,fill=GenericItem,draw] (preprocessor) at (6,1)
                {\textbf{Preprocessor}\\
                 Mapping $\rightarrow$ mapping\\
                 Modify a mapping on the fly};
          \node [text width=5cm,fill=GenericItem,draw] (generator) at (6,-1)
                {\textbf{Generator}\\
                 Mapping $\rightarrow$ OWL axiom\\
                 Derive an axiom from a mapping};
          \draw[->] (filter) -- (preprocessor);
          \draw[->] (filter) -- (generator);
        \end{tikzpicture}
      \end{center}
    \end{block}
  }
  \only<2>{
    \begin{block}{Inverting a mapping}\ttfamily\scriptsize
subject==UBERON:* -> invert();
    \end{block}

    \begin{block}{Generating a simple axiom}\ttfamily\scriptsize
predicate==skos:broadMatch -> create\_axiom("\%subject\_id SubClassOf: \%object\_id");
    \end{block}

    \begin{block}{Generating a more complex axiom}\ttfamily\scriptsize
subject==FBbt:*\\
\hspace{1cm} \&\& predicate==semapv:crossSpeciesExactMatch\\
\hspace{1cm} \&\& (mapping\_justification==semapv:ManualMappingCuration || confidence>=0.8)\\
->  create\_axiom("\%subject\_id EquivalentTo: \%object\_id and\\
\hspace{2cm} (BFO:0000050 some NCBITaxon:7227"));
    \end{block}
  }
\end{frame}

\begin{frame}
  \frametitle{Implementation in ROBOT commands}

  \begin{block}{sssom:xref-extract command}
    Convert existing mappings represented as cross-references into a SSSOM set

    \medskip
    {\ttfamily\scriptsize
robot sssom:xref-extract -i input.owl --mapping-file output.sssom.tsv
    }
  \end{block}

  \begin{block}{sssom:inject command}
    Apply SSSOM/T-OWL rules to a mapping set and inject the resulting axioms into an ontology

    \medskip
    {\ttfamily\scriptsize
robot sssom:inject -i input.owl --sssom mappings.sssom.tsv --ruleset bridge.rules
    }
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Integrating the species-specific ontologies}

  \begin{center}
    \begin{tikzpicture}[node distance=.5cm]
      \node [Entity] (uberon) {Uberon};
      \node [Entity,right=of uberon] (com) {Mapping sets};
      \node [Entity,right=of com] (ssao) {Taxon-specific ontologies};
      \node [Process,below=of com] (sssomt) {SSSOM/T-OWL};
      \node [Entity,below=of sssomt] (cob) {Bridge files};
      \node [Process,below=of cob] (merge) {OWL Merge};
      \node [Entity,below=of merge] (cm) {Multi-species ontology};
      \draw [->] (com) -- (sssomt) (sssomt) -- (cob);
      \draw [->] (cob) -- (merge) (merge) -- (cm);
      \draw [->] (uberon.south) |- (merge.west);
      \draw [->] (ssao.south) |- (merge.east);
    \end{tikzpicture}
  \end{center}
\end{frame}

\section{Conclusion}

\begin{frame}
  \frametitle{Outcome}

  \begin{block}{Cross-ontology mappings as first-class artefacts}
    \begin{itemize}
      \item With their own metadata
      \item Can be maintained separately
      \item Available as a separate release product in a standard format
    \end{itemize}
  \end{block}

  \begin{block}{Generic framework for SSSOM-based integration of multiple ontologies}
    \begin{itemize}
      \item SSSOM/T-OWL is completely independent of Uberon
      \item Only the ruleset is specific
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Example: FBbt/Uberon/CL integration test}

  A single ROBOT command to merge FBbt with Uberon, CL, and the mappings-derived
  bridging axioms, and check for possible inconsistencies that would reveal
  incorrect mappings:

  \begin{block}{}
\begin{verbatim}
robot merge -i fbbt.owl -i uberon.owl -i cl.owl \
      sssom:inject --sssom mappings.sssom.tsv   \
                   --ruleset bridging.rules     \
      reason -r ELK
\end{verbatim}
  \end{block}
\end{frame}

\end{document}